---
title: "Leontief Input-Output matrix"
author: "John Karuitha"
date: "4/29/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Leontief Input-Output Matrix
### Introduction
- Used in economic planning. 

- Especially in centrally planned economies.

- Vasilly Leotief. 

### The concept

- Suppose you produce 100 kgs of maize, and consume half of it. 
- How much can you sell? 100 - 0.5(100) = 50
- Suppose you produce X kgs and consume 0.7 (a fraction of X). 
- How much can you sell? X - 0.7X
- Suppose you produce X kgs and consume A (A is a fraction of X). 
- How much can you sell? X - AX = X(1 - A).

- What is the equivalent of ZERO in matrix algebra?
    - If you want to show that there is nothing in matrix algebra, how would you           write it? It has only Zeros. NULL matrix. It is a matrix that has zeros as           elements. 
- What is the equivalent of ONE in matrix algebra? The IDENTITY matrix. It has ones i the main diagonal, and zeros everywhere else. 

```{r}
## Example of a NULL matrix
matrix(c(0,0,0,0,0,0,0,0,0), nrow = 3)

## Example of an identity matrix
matrix(c(1,0,0,0,1,0,0,0,1), byrow= TRUE, nrow = 3)
```

- What is the symbol for Identity matrix? I

- X(I - A) ### NB: This represents a surplus or excess produce that you can sell out. 
- We call this the surplus or the final demand, D. 

- Hence we can write, D = X(I - A), where A is the amount consumed internally. 

- Let us try to solve for X. 

- $X = \frac{D}{(I - A)} = \frac{1}{(I - A)} * D = (I - A)^{-1} * D$

$X = (I - A)^{-1} * D$

- NB: X is what you need to produce. 

- A is the fraction of the produce you use internally. 

- D is the final demand or surplus. The amount you can sell outside after satisfying your internal demand. 

- Given A and D, can you solve for X? This is the central idea of the Leontief input-Output matrix. 

## Examples in the slides: 


```{r}
I <- matrix(c(1,0,0,0,1,0,0,0,1), byrow= TRUE, nrow = 3)
I

A <- matrix(c(0.5,0.1,0.1,0.2,0.5,0.3,0.1,0.3,0.4), byrow= TRUE, nrow = 3)
A
```

```{r}
W = I - A
W

W_inverse <- solve(W)

D = matrix(c(85,65,0), byrow = FALSE, nrow = 3)
D

W_inverse %*% D
```

